---
title: "Shiny Background"
author: "Tom Pham"
date: "5/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What is Shiny?
Shiny is a framework for creating web applications using R code. <br>


## Shiny Basic Concepts

There are two fundamental ways to create a Shiny app. The first is to have a 
single file called app.R. This file will tell Shiny how your app should look
and how it should behave. <br>


The second way is to to split up the app into 2 files. ui.R and server.R. What
this does is separate the code that defines how you want the app to look (ui.R)
from the code that defines how you want the app to behave (server.R). For my
telemetry app, I've chosen to go with the former. <br>


A minimal example of an app might look like this: <br>
```{r, eval=FALSE}
library(shiny)
ui <- fluidPage(
  "Hello, world!"
)
server <- function(input, output, session) {
}
shinyApp(ui, server)
```

## UI Basics
The user interface is where you define the layout of the app

Here is a list of all possible UI Layout functions:

